{\steve}: A hierarchical Bayesian model for Supernova Cosmology

### 2018-10-14 papers/methods/paper.tex
Sooo finally got emulatapj working, now I need to rejig all the tables

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Basic setup. Most papers should leave these options alone.
\documentclass[a4paper,fleqn,usenatbib]{emulateapj}

@@

% Only include extra packages if you really need them. Common packages are:
\usepackage{graphicx}	% Including figure files
\usepackage{amsmath}	% Advanced maths commands
\usepackage{amssymb}	% Extra maths symbols
\usepackage{tikz}
\usepackage{enumerate}
%\usepackage{tabu}
%\usepackage{tabularx}
\DeclareMathAlphabet{\mathcal}{OMS}{cmsy}{m}{n}
\usetikzlibrary{bayesnet}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\shorttitle{\name}
% Don't change these lines
\begin{document}
\title{\name: A hierarchical Bayesian model for Supernova Cosmology}

 The list of authors, and the short list which is used in the headers.
 % If you need two or more lines of authors, add an extra line using \newauthor
 \author{
 	S.~R.~Hinton\altaffilmark{1,2},\thanks{E-mail: samuelreay@gmail.com}
 	T.~M.~Davis\altaffilmark{1,2},
 	A.~G.~Kim\altaffilmark{3},
 	A.~M\"oller\altaffilmark{2,4},
 	M.~Sako\altaffilmark{5},
 	M.~Smith\altaffilmark{6},
 	\\}
 \affil{$^{1}$ School of Mathematics and Physics, University of Queensland,  Brisbane, QLD 4072, Australia}
 \affil{$^{2}$ ARC Centre of Excellence for All-sky Astrophysics (CAASTRO)}
 \affil{$^{3}$ Lawrence Berkeley National Laboratory, 1 Cyclotron Road, Berkeley, CA 94720, USA}
 \affil{$^{4}$ The Research School of Astronomy and Astrophysics, Australian National University, ACT 2601, Australia}
 \affil{$^{5}$ Department of Physics and Astronomy, University of Pennsylvania, Philadelphia, PA 19104, USA}
 \affil{$^{6}$ School of Physics and Astronomy, University of Southampton,  Southampton, SO17 1BJ, UK}

@@

@@

\begin{align}
\mu_{\rm obs} = m_B + \alpha x_1 - \beta c - M_B + \Delta M \cdot m + \text{other corrections}, \label{eq:standard}
\end{align}

@@

\begin{align}
\chi^2 = (\mu_{\rm obs} - \mu_C)^\dagger C^{-1}_{\mu} (\mu_{\rm obs} - \mu_C),
\end{align}

@@

\begin{align}
\mu_C &= 5 \log\left[ \frac{d_L}{10{\rm pc}} \right], \label{eq:mu0}\\
d_L& = (1+z)\frac{c}{H_0} \int_0^z \frac{dz'}{E(z')},\\
E(z) &= \sqrt{ \Omega_m (1+z')^3 + \Omega_k (1 + z')^2 + \Omega_\Lambda (1+z')^{3(1+w)}}
\end{align}

@@

For the moment, let us consider a single supernova and its classic summary statistics $m_B$, $x_1$, $c$. For convenience, let us define $\eta \equiv \lbrace m_B, x_1, c \rbrace$. A full treatment of the summary statistics would involve determining $p(\hat{y}|\eta)$, where $\hat{y}$ represents the observed light curves fluxes and uncertainties. However, this is computationally prohibitive  as it would require incorporating SALT2 light curve fitting inside our model fitting. Due to this computational expense, we rely on initially fitting the light curve observations to produce a best fit $\hat{\eta}$ along with a $3\times3$ covariance matrix $\mathcal{C_\eta}$ describing the uncertainty on $\hat{\eta}$. Using this simplification, our latent variables are given by
\begin{align}
p(\hat{\eta}|\eta) \sim \mathcal{N}(\hat{\eta} | \eta, \mathcal{C_\eta}). \label{eq:pop}
\end{align}
As discussed in Section \ref{sec:challenges},  the SALT2 model does not include full treatment of intrinsic dispersion, and thus this approximation does not encapsulate the full uncertainty introduced from this dispersion.

% \begin{figure}
% 	\begin{tikzpicture}
% 	% Obs nodes
% 	\node[obs]  (obssum) {$\hat{m}_B, \hat{x}_1, \hat{c}, \mathcal{C_\eta}^{-1}$};
% 	\node[obs, left=0.8cm of obssum]  (zobs) {$\hat{z}$};
% 	\node[obs, right=0.8cm of obssum]  (mobs) {$\hat{m}$};
% 	\node[obs, right=0.8cm of mobs]  (pobs) {$\hat{p}$};
% 	% Latent nodes
% 	\node[latent, above=of obssum] (sum) {$m_B, x_1, c$};
% 	\node[det, above=of zobs] (z) {$z$};
% 	\node[det, above=of z] (mu) {$\mu$};
% 	\node[latent, right=4.33cm of mu] (MBobs) {$M_B$};
% 	% Survey nodes
% 	\node[latent, above=of sum, yshift=1cm,text width=2.7cm, align=center] (pop1) {$\langle x_1^i \rangle, \langle c^i \rangle,\sigma_{M_B}, $\\$ \sigma_{x_1}, \sigma_c, \alpha_c, \kappa_0, \kappa_1$ };
% 	\node[latent, right=1.5cm of pop1, align=center] (sf) {$\delta S$};
% 	% Global nodes
% 	\node[latent, above=2cm of mu]  (cosmo) {$\Omega_m, w$};
% 	\node[latent, above=1.7cm of MBobs, text width=3cm, align=center]  (ab) {$\langle M_B \rangle, \alpha, \beta$, $\delta(0)$, \\$\delta(\infty)$, $\delta \mathcal{Z}_i$};
% 	% Connect the nodes
% 	\edge {sum} {obssum};
% 	\edge {z} {zobs};
% 	\edge {z,cosmo} {mu};
% 	\edge {mu, sum, mobs, pop1, ab, pobs, sf} {MBobs};
% 	% Plates
% 	\plated[thick] {sn} {(obssum)(zobs)(mobs)(pobs)(sum)(z)(mu)} {$\forall\ \rm{SN}$} ;
% 	\plated[thick] {survey} {(sn)(pop1)} {\\$\forall\ \rm{Survey}$} ;
% 	\end{tikzpicture}
% 	\caption{Probabilistic graphical model for our likelihood. Double-lined nodes represent observed variables, diamond nodes represent deterministic variables, single-lined ellipse nodes represent fit variables. The SN box represents observed variables and latent variables for each individual supernova, whilst the survey box represents survey-specific variables, which in general describe the supernova population for the survey and the systematics associated with it. Variables that appear outside both boxes represent top level model parameters.}
% 	\label{fig:pgm}
% \end{figure}

@@

\begin{align}
P(M_B, x_1, c, z | \langle M_B \rangle, \langle x_1(z) \rangle, \langle c(z) \rangle, \sigma_{m_B}, \sigma_{x_1}, \sigma_c, \alpha_c) = \notag \\
\mathcal{N}(M_B|\langle M_B \rangle, \sigma_{M_B}) \mathcal{N}(x_1 | \langle x_1(z) \rangle, \sigma_{x_1}) \mathcal{N}^{\rm{skew}}(c| \langle c(z) \rangle, \sigma_c, \alpha_c) \label{eq:l1}.
\end{align}

@@

\begin{align}
\Delta M = \delta(0) \left[ \frac{1.9\left(1 - \frac{\delta(0)}{\delta(\infty)}\right)  }{0.9 + 10^{0.95z}} + \frac{\delta(0)}{\delta(\infty)}\right], \label{eq:mass}
\end{align}

@@

\begin{align}
P\left(\hat{\eta}, \frac{\partial \hat{\eta}}{\partial \Z_i} | \eta, \delta \Z_i, C\right) = \mathcal{N}\left(\hat{\eta} + \delta \Z_i \frac{\partial \hat{\eta}}{\partial \Z_i}|\eta,\cov\right). \label{eq:l3}
\end{align}

@@

\begin{align}
d &= \int  \left[ \int P(S|m_B) P(m_B | z, \theta)\, d m_B \right] P(S|z) P(z|\theta)\, dz, \label{eq:w1}
\end{align}

@@

\begin{align}
m_B^*(z) &= \langle M_B \rangle + \mu(z) - \alpha \langle x_1(z) \rangle + \beta \langle c(z) \rangle \\
\sigma^{*2}_{m_B} &= \sigma_{M_B}^2 + (\alpha \sigma_{x_1})^2 + (\beta \sigma_c)^2.
\end{align}

@@

\begin{align}
P(S|m_B) = \Phi^c(m_B | \mu_{\rm CDF}, \sigma_{\rm CDF}),
\end{align}

@@

\begin{align}
P(S|m_B) = \mathcal{N}^{{\rm Skew}}(m_B | \mu_{\rm Skew}, \sigma_{\rm Skew}, \alpha_{\rm Skew}).
\end{align}

@@

\begin{align}
d_{\rm CDF} &= \Phi^c\left(  \frac{m^*_B - \mu_{\rm CDF}}{\sqrt{\sigma^{*2}_{m_B} + \sigma_{\rm CDF}^2}}  \right) \label{eq:seldes}\\
d_{\rm Skew} &= 2\mathcal{N}\left( \frac{m_B^* - \mu_{\rm Skew}}{\sqrt{\sigma^{*2}_{m_B} + \sigma_{\rm Skew}^2}}\right) \times \notag\\
&\quad\quad \Phi\left(  \frac{{\rm sign}(\alpha_{\rm Skew})(m_B^* - \mu_{\rm Skew})}{\frac{\sigma_{m_B}^{*2} + \sigma^2_{\rm Skew}}{\sigma^2_{\rm Skew}} \sqrt{\frac{\sigma^2_{\rm Skew}}{\alpha^2_{\rm Skew}} + \frac{\sigma_{m_B}^{*2} \sigma^2_{\rm Skew}}{\sigma_{m_B}^{*2} + \sigma^2_{\rm Skew}}} }\right), \label{eq:sellow-$z$}
\end{align}

@@

\subsection{General Selection Effects}
\label{app:selection1}
When formulating and fitting a model using a constraining dataset, we wish to resolve the posterior surface defined by
\begin{align}
P(\theta | {\rm data}) \propto P({\rm data} | \theta) P(\theta),
\end{align}
which gives the probability of the model parameter values ($\theta$) given the data.  Prior knowledge of the allowed values of the model parameters is encapsulated in the prior probability $P(\theta)$. Of primary interest to us is the likelihood of observing the data given our parametrised model, $\mathcal{L} \equiv P({\rm data} | \theta)$. When dealing with experiments that have imperfect selection efficiency, our likelihood must take that efficiency into account.  We need to describe the probability that the events we observe are both drawn from the distribution predicted by the underlying theoretical model \textit{and} that those events, given they happened, are subsequently successfully observed.  To make this extra conditional explicit, we write the likelihood of the data given an underlying model, $\theta$, \textit{and} that the data are included in our sample, denoted by $S$, as:
\begin{align}
\mathcal{L} &= P({\rm data} | \theta, S). \label{eq:app_like}
\end{align}
A variety of selection criteria are possible, and in our method we use our data in combination with the proposed model to determine the probability of particular selection criteria.  That is, we characterise a function $P(S|{\rm data},\theta)$, which colloquially can be stated as \textit{the probability of a potential observation passing selection cuts, given our observations and the underlying model}. We can introduce this expression in a few lines due to symmetry of joint probabilities and utilising that $P(x,y,z) = P(x|y,z)P(y,z) = P(y|x, z)P(x, z)$:
\begin{align}
P({\rm data} | S , \theta) P(S, \theta) &= P(S | {\rm data}, \theta) P({\rm data}, \theta)\\
P({\rm data} | S , \theta) &= \frac{ P(S | {\rm data}, \theta) P({\rm data}, \theta) }{ P(S, \theta) }\\
&= \frac{ P(S | {\rm data}, \theta) P({\rm data} | \theta) P(\theta)}{ P(S | \theta)  P(\theta)}\\
&= \frac{ P(S | {\rm data}, \theta) P({\rm data} | \theta)}{ P(S | \theta) }
\end{align}
which is equal to the likelihood $\mathcal{L}$. Introducing an integral over all possible events $D$, so we can evaluate $P(S|\theta)$, 
\begin{align}
\mathcal{L}&= \frac{P(S|{\rm data},\theta) P({\rm data}|\theta)}{\int P(S, D|\theta)\, dD} \\
\mathcal{L} &= \frac{P(S|{\rm data},\theta) P({\rm data}|\theta)}{\int P(S | D, \theta) P(D|\theta)\, dD}, \label{eq:app_main}
\end{align}
where we define the denominator as $d$ for simplicity in future derivations.

\subsection{Supernova Selection Effects}
\label{app:selection2}
We assume that our selection effects can be reasonably well encapsulated by independent functions of (true) apparent magnitude and redshift, such that $P(S|{\rm data}, \theta) = P(S|z) P(S|m_B)$. Our denominator then becomes
\begin{align}
d &= \int d\hat{z} \, d\hat{m}_B \, dz \, dm_B P(S|z) P(S|m_B) P(\hat{z}|z) P(\hat{m}_B|m_B) P(z, m_B | \theta), \label{eq:sel}
\end{align}
where for simplicity we have not written out all the integrals that do not interact with the selection effects explicitly. Due to our assumed perfect measurement of redshift, $P(\hat{z}|z) = \delta(\hat{z} - z)$. $P(\hat{m}_B | m_B)$ is a Gaussian due to our Gaussian model of summary statistics, and $m_B$, $x_1$, $c$, and can be analytically integrated out, collapsing the integral over $\hat{m}_B$ (which is why they were not included in equation \eqref{eq:sel}). Finally, we can express $P(z, m_B | \theta)$ as  $P(m_B | z, \theta) P(z | \theta)$, where the first term requires us to calculate the magnitude distribution of our underlying population at a given redshift, and the second term is dependent on survey geometry and supernovae rates. We can thus state
\begin{align}
d = \int \left[ \int P(S|m_B) P(m_B | z, \theta)\, dm_B \right] P(S|z)P(z|\theta)\, dz.
\end{align}
By assuming that the distribution $P(S|z)P(z|\theta)$ is well sampled by the observed supernova redshifts, we can approximate the integral over redshift by evaluating
\begin{align}
\int P(S|m_B) P(m_B | z, \theta)\, dm_B \label{eq:selint}
\end{align}
for each supernova in the dataset -- i.e. Monte Carlo integration with assumed perfect importance sampling.

As stated in Section \ref{sec:selection}, the underlying population in apparent magnitude, when we discard skewness, can be represented as $\mathcal{N}(m_B|m_B^*(z), \sigma^*_{m_B})$, where
\begin{align}
m_B^*(z) &= \langle M_B \rangle + \mu(z) - \alpha \langle x_1(z) \rangle + \beta \left(\langle c(z) \rangle + \sqrt{\frac{2}{\pi}}\sigma_c \delta_c\right)\label{eq:disc1} \\
\sigma^*_{m_B} &= \sigma_{M_B}^2 + (\alpha \sigma_{x_1})^2 +  \left(\beta \sigma_c \sqrt{1 - \frac{2\delta_c^2}{\pi}}\right)^2. \label{eq:disc2}
\end{align}
Then, modelling $P(S|m_B)$ as either a normal or a skew normal, we can analytically perform the integral in equation \eqref{eq:selint} and reach equations \eqref{eq:seldes} and \eqref{eq:sellow-$z$}.

@@

In this section, we investigate the effect of approximating the skew normal underlying color distribution as a normal. Specifically, equations \eqref{eq:disc1} and \eqref{eq:disc2} make the assumption that, for our color distribution, $\mathcal{N}^{\rm Skew}(\mu, \sigma, \alpha)$ is well approximated by $\mathcal{N}(\mu, \sigma)$. We sought to improve on this approximation by adjusting the mean and standard deviation of the approximated normal to more accurately describe the actual mean and standard deviation of a skew normal. With $\delta \equiv \alpha/\sqrt{1+\alpha^2}$, the correct mean and standard deviation are
\begin{align}
\mu_1 &= \mu_0 + \sqrt{\frac{2}{\pi}} \delta \sigma_0 \\
\sigma_1 &= \sigma_0 \sqrt{1 - \frac{2 \delta^2}{\pi}},
\end{align}
where we highlight that $\mu$ here represents the mean of the distribution, not distance modulus. We can then test the approximation $\mathcal{N}^{\rm Skew}(\mu_0, \sigma_0, \alpha) \rightarrow \mathcal{N}(\mu_1, \sigma_1)$. Unfortunately, this shift to the mean and standard deviation of the normal approximation where we treat $m_B$, $x_1$, and $c$ as a multivariate skew normal did not produce stable posterior surfaces. Due to this, we treat the underlying $m_B$, $x_1$, and $c$ populations as independent.

@@

Considering a $3\times3$ matrix with cholesky decomposition
\begin{align}
L = \begin{pmatrix}
a & 0 & 0 \\ b & c & 0 \\ d & e & f \\
\end{pmatrix},
\end{align}
the original covariance matrix $\Sigma$ is given by
\begin{align}
\Sigma = \begin{pmatrix}
a^2 & ab & ad \\ ab & b^2 + c^2 & bd + ce \\ ad & bd + ce & d^2 + e^2 + f^2\\
\end{pmatrix}.
\end{align}
Now, the primary source of extra uncertainty in the intrinsic dispersion models comes from chromatic smearing, which primarily influences the recovered color parameter, which is placed as the last element in the observables vector $\lbrace m_B, x_1, c\rbrace$. We can now see that it is possible to add extra uncertainty to the color observation on the diagonal without having to recompute the cholesky decomposition - notice that $f$ is unique in that it is the only element of $L$ that appears in only one position in the covariance matrix. To take our covariance and add on the diagonal uncertainty for color an extra $\sigma_e$ term, we get
\begin{align}
C = \begin{pmatrix}
\sigma_{m_B}^2 & \rho_{0,1} \sigma_{m_B} \sigma_{x_1} & \rho_{0,2} \sigma_{m_B} \sigma_c \\
\rho_{0,1} \sigma_{m_B} \sigma_{x_1} & \sigma_{x_1}^2 & \rho_{1, 2} \sigma_{x_1} \sigma_c \\
\rho_{0,2} \sigma_{m_B} \sigma_c & \rho_{1, 2} \sigma_{x_1} \sigma_c &  \sigma_c^2 + \sigma_e^2 \\
\end{pmatrix}.
\end{align}
The cholesky decomposition of this is, in terms of the original cholesky decomposition, is
\begin{align}
L = \begin{pmatrix}
a & 0 & 0 \\ b & c & 0 \\ d & e & f + g \\
\end{pmatrix},
\end{align}
where $g = \sqrt{f^2 + \sigma_e^2} - f$. This allows an easy update to the cholesky decomposition to add extra uncertainty to the independent color uncertainty. For both the {\gten} and {\celeven} models, we ran fits without the cholesky parametrisation to allow for extra correlated dispersion (instead of just dispersion on $c$), but find no decrease in bias or improved fit statistics, allowing us to use the more efficient cholesky parametrisation.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Basic setup. Most papers should leave these options alone.
\documentclass[a4paper,fleqn,usenatbib,manuscript]{emulateapj}

@@

% Only include extra packages if you really need them. Common packages are:
\usepackage{graphicx}	% Including figure files
\usepackage{amsmath}	% Advanced maths commands
\usepackage{amssymb}	% Extra maths symbols
\usepackage{enumerate}
\usepackage{xcolor}

%\usepackage{tabu}
%\usepackage{tabularx}
\DeclareMathAlphabet{\mathcal}{OMS}{cmsy}{m}{n}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\shorttitle{{\steve}}
% Don't change these lines
\begin{document}
\title{{\steve}: A hierarchical Bayesian model for Supernova Cosmology}

 % If you need two or more lines of authors, add an extra line using \newauthor
 \author{
 	S.~R.~Hinton\altaffilmark{1,2},\thanks{E-mail: samuelreay@gmail.com}
 	T.~M.~Davis\altaffilmark{1,2},
 	A.~G.~Kim\altaffilmark{3},
 	A.~M\"oller\altaffilmark{2,4},
 	M.~Sako\altaffilmark{5},
 	M.~Smith\altaffilmark{6},
 	\\}
 \affil{$^{1}$ School of Mathematics and Physics, University of Queensland,  Brisbane, QLD 4072, Australia}
 \affil{$^{2}$ ARC Centre of Excellence for All-sky Astrophysics (CAASTRO)}
 \affil{$^{3}$ Lawrence Berkeley National Laboratory, 1 Cyclotron Road, Berkeley, CA 94720, USA}
 \affil{$^{4}$ The Research School of Astronomy and Astrophysics, Australian National University, ACT 2601, Australia}
 \affil{$^{5}$ Department of Physics and Astronomy, University of Pennsylvania, Philadelphia, PA 19104, USA}
 \affil{$^{6}$ School of Physics and Astronomy, University of Southampton,  Southampton, SO17 1BJ, UK}

@@

\keywords{cosmology: supernovae}

@@

\begin{equation}
\mu_{\rm obs} = m_B + \alpha x_1 - \beta c - M_B + \Delta M \cdot m + \text{other corrections}, \label{eq:standard}
\end{equation}

@@

\begin{equation}
\chi^2 = (\mu_{\rm obs} - \mu_C)^\dagger C^{-1}_{\mu} (\mu_{\rm obs} - \mu_C),
\end{equation}

@@

\begin{eqnarray}
\mu_C &= 5 \log\left[ \frac{d_L}{10{\rm pc}} \right], \label{eq:mu0}\\
d_L& = (1+z)\frac{c}{H_0} \int_0^z \frac{dz'}{E(z')},\\
E(z) &= \sqrt{ \Omega_m (1+z')^3 + \Omega_k (1 + z')^2 + \Omega_\Lambda (1+z')^{3(1+w)}}
\end{eqnarray}

@@

For the moment, let us consider a single supernova and its classic summary statistics $m_B$, $x_1$, $c$. For convenience, let us define $\eta \equiv \lbrace m_B, x_1, c \rbrace$. A full treatment of the summary statistics would involve determining $p(\hat{y}|\eta)$, where $\hat{y}$ represents the observed light curves fluxes and uncertainties. However, this is computationally prohibitive  as it would require incorporating SALT2 light curve fitting inside our model fitting. Due to this computational expense, we rely on initially fitting the light curve observations to produce a best fit $\hat{\eta}$ along with a $3\times3$ covariance matrix $\mathcal{C_\eta}$ describing the uncertainty on $\hat{\eta}$. Using this simplification, our latent variables are given by
\begin{equation}
p(\hat{\eta}|\eta) \sim \mathcal{N}(\hat{\eta} | \eta, \mathcal{C_\eta}). \label{eq:pop}
\end{equation}
As discussed in Section \ref{sec:challenges},  the SALT2 model does not include full treatment of intrinsic dispersion, and thus this approximation does not encapsulate the full uncertainty introduced from this dispersion.

 \begin{figure}
 	\begin{center}
 		\includegraphics[width=\columnwidth]{tikz-figure0.pdf}
 	\end{center}
 	\caption{Probabilistic graphical model for our likelihood. Double-lined nodes represent observed variables, diamond nodes represent deterministic variables, single-lined ellipse nodes represent fit variables. The SN box represents observed variables and latent variables for each individual supernova, whilst the survey box represents survey-specific variables, which in general describe the supernova population for the survey and the systematics associated with it. Variables that appear outside both boxes represent top level model parameters.}
 	\label{fig:pgm}
 \end{figure}

@@

\begin{equation}
P(M_B, x_1, c, z | \langle M_B \rangle, \langle x_1(z) \rangle, \langle c(z) \rangle, \sigma_{m_B}, \sigma_{x_1}, \sigma_c, \alpha_c) = \notag \\
\mathcal{N}(M_B|\langle M_B \rangle, \sigma_{M_B}) \mathcal{N}(x_1 | \langle x_1(z) \rangle, \sigma_{x_1}) \mathcal{N}^{\rm{skew}}(c| \langle c(z) \rangle, \sigma_c, \alpha_c) \label{eq:l1}.
\end{equation}

@@

\begin{equation}
\Delta M = \delta(0) \left[ \frac{1.9\left(1 - \frac{\delta(0)}{\delta(\infty)}\right)  }{0.9 + 10^{0.95z}} + \frac{\delta(0)}{\delta(\infty)}\right], \label{eq:mass}
\end{equation}

@@

\begin{equation}
P\left(\hat{\eta}, \frac{\partial \hat{\eta}}{\partial \Z_i} | \eta, \delta \Z_i, C\right) = \mathcal{N}\left(\hat{\eta} + \delta \Z_i \frac{\partial \hat{\eta}}{\partial \Z_i}|\eta,\cov\right). \label{eq:l3}
\end{equation}

@@

\begin{equation}
d = \int  \left[ \int P(S|m_B) P(m_B | z, \theta)\, d m_B \right] P(S|z) P(z|\theta)\, dz, \label{eq:w1}
\end{equation}

@@

\begin{eqnarray}
m_B^*(z) &= \langle M_B \rangle + \mu(z) - \alpha \langle x_1(z) \rangle + \beta \langle c(z) \rangle \\
\sigma^{*2}_{m_B} &= \sigma_{M_B}^2 + (\alpha \sigma_{x_1})^2 + (\beta \sigma_c)^2.
\end{eqnarray}

@@

\begin{equation}
P(S|m_B) = \Phi^c(m_B | \mu_{\rm CDF}, \sigma_{\rm CDF}),
\end{equation}

@@

\begin{equation}
P(S|m_B) = \mathcal{N}^{{\rm Skew}}(m_B | \mu_{\rm Skew}, \sigma_{\rm Skew}, \alpha_{\rm Skew}).
\end{equation}

@@

\begin{eqnarray}
d_{\rm CDF} &= \Phi^c\left(  \frac{m^*_B - \mu_{\rm CDF}}{\sqrt{\sigma^{*2}_{m_B} + \sigma_{\rm CDF}^2}}  \right) \label{eq:seldes}\\
d_{\rm Skew} &= 2\mathcal{N}\left( \frac{m_B^* - \mu_{\rm Skew}}{\sqrt{\sigma^{*2}_{m_B} + \sigma_{\rm Skew}^2}}\right) \times \notag\\
&\quad\quad \Phi\left(  \frac{{\rm sign}(\alpha_{\rm Skew})(m_B^* - \mu_{\rm Skew})}{\frac{\sigma_{m_B}^{*2} + \sigma^2_{\rm Skew}}{\sigma^2_{\rm Skew}} \sqrt{\frac{\sigma^2_{\rm Skew}}{\alpha^2_{\rm Skew}} + \frac{\sigma_{m_B}^{*2} \sigma^2_{\rm Skew}}{\sigma_{m_B}^{*2} + \sigma^2_{\rm Skew}}} }\right), \label{eq:sellow-$z$}
\end{eqnarray}

@@

\subsection{General Selection Effects}
\label{app:selection1}
When formulating and fitting a model using a constraining dataset, we wish to resolve the posterior surface defined by
\begin{equation}
P(\theta | {\rm data}) \propto P({\rm data} | \theta) P(\theta),
\end{equation}
which gives the probability of the model parameter values ($\theta$) given the data.  Prior knowledge of the allowed values of the model parameters is encapsulated in the prior probability $P(\theta)$. Of primary interest to us is the likelihood of observing the data given our parametrised model, $\mathcal{L} \equiv P({\rm data} | \theta)$. When dealing with experiments that have imperfect selection efficiency, our likelihood must take that efficiency into account.  We need to describe the probability that the events we observe are both drawn from the distribution predicted by the underlying theoretical model \textit{and} that those events, given they happened, are subsequently successfully observed.  To make this extra conditional explicit, we write the likelihood of the data given an underlying model, $\theta$, \textit{and} that the data are included in our sample, denoted by $S$, as:
\begin{equation}
\mathcal{L} = P({\rm data} | \theta, S). \label{eq:app_like}
\end{equation}
A variety of selection criteria are possible, and in our method we use our data in combination with the proposed model to determine the probability of particular selection criteria.  That is, we characterise a function $P(S|{\rm data},\theta)$, which colloquially can be stated as \textit{the probability of a potential observation passing selection cuts, given our observations and the underlying model}. We can introduce this expression in a few lines due to symmetry of joint probabilities and utilising that $P(x,y,z) = P(x|y,z)P(y,z) = P(y|x, z)P(x, z)$:
\begin{eqnarray}
P({\rm data} | S , \theta) P(S, \theta) &= P(S | {\rm data}, \theta) P({\rm data}, \theta)\\
P({\rm data} | S , \theta) &= \frac{ P(S | {\rm data}, \theta) P({\rm data}, \theta) }{ P(S, \theta) }\\
&= \frac{ P(S | {\rm data}, \theta) P({\rm data} | \theta) P(\theta)}{ P(S | \theta)  P(\theta)}\\
&= \frac{ P(S | {\rm data}, \theta) P({\rm data} | \theta)}{ P(S | \theta) }
\end{eqnarray}
which is equal to the likelihood $\mathcal{L}$. Introducing an integral over all possible events $D$, so we can evaluate $P(S|\theta)$, 
\begin{eqnarray}
\mathcal{L}&= \frac{P(S|{\rm data},\theta) P({\rm data}|\theta)}{\int P(S, D|\theta)\, dD} \\
\mathcal{L} &= \frac{P(S|{\rm data},\theta) P({\rm data}|\theta)}{\int P(S | D, \theta) P(D|\theta)\, dD}, \label{eq:app_main}
\end{eqnarray}
where we define the denominator as $d$ for simplicity in future derivations.

\subsection{Supernova Selection Effects}
\label{app:selection2}
We assume that our selection effects can be reasonably well encapsulated by independent functions of (true) apparent magnitude and redshift, such that $P(S|{\rm data}, \theta) = P(S|z) P(S|m_B)$. Our denominator then becomes
\begin{equation}
d = \int d\hat{z} \, d\hat{m}_B \, dz \, dm_B P(S|z) P(S|m_B) P(\hat{z}|z) P(\hat{m}_B|m_B) P(z, m_B | \theta), \label{eq:sel}
\end{equation}
where for simplicity we have not written out all the integrals that do not interact with the selection effects explicitly. Due to our assumed perfect measurement of redshift, $P(\hat{z}|z) = \delta(\hat{z} - z)$. $P(\hat{m}_B | m_B)$ is a Gaussian due to our Gaussian model of summary statistics, and $m_B$, $x_1$, $c$, and can be analytically integrated out, collapsing the integral over $\hat{m}_B$ (which is why they were not included in equation \eqref{eq:sel}). Finally, we can express $P(z, m_B | \theta)$ as  $P(m_B | z, \theta) P(z | \theta)$, where the first term requires us to calculate the magnitude distribution of our underlying population at a given redshift, and the second term is dependent on survey geometry and supernovae rates. We can thus state
\begin{equation}
d = \int \left[ \int P(S|m_B) P(m_B | z, \theta)\, dm_B \right] P(S|z)P(z|\theta)\, dz.
\end{equation}
By assuming that the distribution $P(S|z)P(z|\theta)$ is well sampled by the observed supernova redshifts, we can approximate the integral over redshift by evaluating
\begin{equation}
\int P(S|m_B) P(m_B | z, \theta)\, dm_B \label{eq:selint}
\end{equation}
for each supernova in the dataset -- i.e. Monte Carlo integration with assumed perfect importance sampling.

As stated in Section \ref{sec:selection}, the underlying population in apparent magnitude, when we discard skewness, can be represented as $\mathcal{N}(m_B|m_B^*(z), \sigma^*_{m_B})$, where
\begin{eqnarray}
m_B^*(z) &= \langle M_B \rangle + \mu(z) - \alpha \langle x_1(z) \rangle + \beta \left(\langle c(z) \rangle + \sqrt{\frac{2}{\pi}}\sigma_c \delta_c\right)\label{eq:disc1} \\
\sigma^*_{m_B} &= \sigma_{M_B}^2 + (\alpha \sigma_{x_1})^2 +  \left(\beta \sigma_c \sqrt{1 - \frac{2\delta_c^2}{\pi}}\right)^2. \label{eq:disc2}
\end{eqnarray}
Then, modelling $P(S|m_B)$ as either a normal or a skew normal, we can analytically perform the integral in equation \eqref{eq:selint} and reach equations \eqref{eq:seldes} and \eqref{eq:sellow-$z$}.

@@

In this section, we investigate the effect of approximating the skew normal underlying color distribution as a normal. Specifically, equations \eqref{eq:disc1} and \eqref{eq:disc2} make the assumption that, for our color distribution, $\mathcal{N}^{\rm Skew}(\mu, \sigma, \alpha)$ is well approximated by $\mathcal{N}(\mu, \sigma)$. We sought to improve on this approximation by adjusting the mean and standard deviation of the approximated normal to more accurately describe the actual mean and standard deviation of a skew normal. With $\delta \equiv \alpha/\sqrt{1+\alpha^2}$, the correct mean and standard deviation are
\begin{eqnarray}
\mu_1 &= \mu_0 + \sqrt{\frac{2}{\pi}} \delta \sigma_0 \\
\sigma_1 &= \sigma_0 \sqrt{1 - \frac{2 \delta^2}{\pi}},
\end{eqnarray}
where we highlight that $\mu$ here represents the mean of the distribution, not distance modulus. We can then test the approximation $\mathcal{N}^{\rm Skew}(\mu_0, \sigma_0, \alpha) \rightarrow \mathcal{N}(\mu_1, \sigma_1)$. Unfortunately, this shift to the mean and standard deviation of the normal approximation where we treat $m_B$, $x_1$, and $c$ as a multivariate skew normal did not produce stable posterior surfaces. Due to this, we treat the underlying $m_B$, $x_1$, and $c$ populations as independent.

@@

Considering a $3\times3$ matrix with cholesky decomposition
\begin{equation}
L = \begin{pmatrix}
a & 0 & 0 \\ b & c & 0 \\ d & e & f \\
\end{pmatrix},
\end{equation}
the original covariance matrix $\Sigma$ is given by
\begin{equation}
\Sigma = \begin{pmatrix}
a^2 & ab & ad \\ ab & b^2 + c^2 & bd + ce \\ ad & bd + ce & d^2 + e^2 + f^2\\
\end{pmatrix}.
\end{equation}
Now, the primary source of extra uncertainty in the intrinsic dispersion models comes from chromatic smearing, which primarily influences the recovered color parameter, which is placed as the last element in the observables vector $\lbrace m_B, x_1, c\rbrace$. We can now see that it is possible to add extra uncertainty to the color observation on the diagonal without having to recompute the cholesky decomposition - notice that $f$ is unique in that it is the only element of $L$ that appears in only one position in the covariance matrix. To take our covariance and add on the diagonal uncertainty for color an extra $\sigma_e$ term, we get
\begin{equation}
C = \begin{pmatrix}
\sigma_{m_B}^2 & \rho_{0,1} \sigma_{m_B} \sigma_{x_1} & \rho_{0,2} \sigma_{m_B} \sigma_c \\
\rho_{0,1} \sigma_{m_B} \sigma_{x_1} & \sigma_{x_1}^2 & \rho_{1, 2} \sigma_{x_1} \sigma_c \\
\rho_{0,2} \sigma_{m_B} \sigma_c & \rho_{1, 2} \sigma_{x_1} \sigma_c &  \sigma_c^2 + \sigma_e^2 \\
\end{pmatrix}.
\end{equation}
The cholesky decomposition of this is, in terms of the original cholesky decomposition, is
\begin{equation}
L = \begin{pmatrix}
a & 0 & 0 \\ b & c & 0 \\ d & e & f + g \\
\end{pmatrix},
\end{equation}
where $g = \sqrt{f^2 + \sigma_e^2} - f$. This allows an easy update to the cholesky decomposition to add extra uncertainty to the independent color uncertainty. For both the {\gten} and {\celeven} models, we ran fits without the cholesky parametrisation to allow for extra correlated dispersion (instead of just dispersion on $c$), but find no decrease in bias or improved fit statistics, allowing us to use the more efficient cholesky parametrisation.
